Pool Model Paper

### 2023-09-01 pool_model.tex
add Temprture stress Figures

One of the reasons of the dormant state appearence can be also the temperature shift~\cite{oliver_viable_1995}.
Moreover, the rapid shift in temperature is associated with the appearence of an additional lag phase~\cite{zwietering_modeling_1994}.
To model it we propose a kind of visco-elastic Maxwell type of stress-strain relation and write down the ordinary differential equation for $\gamma$ to be:
\begin{eqnarray}
\dot{\gamma} &=& \Gamma \left |\dot{T}\right |-\delta \gamma.
\end{eqnarray}
We assumed that the direction of the temperature change does not matter, i.e., a change from $T=2$ to $T=12$ has the same effect as a change from $T=14$ to $T=4$. 
$\Gamma$ is a scaling factor and $\delta$ is the relaxation time for the temperature disturbance. 
Integration of this equation yields (with $\gamma(0)=0$):
\begin{eqnarray}
\gamma(t) &=& \Gamma \int_0^t \left |\dot{T}(t^{\prime})\right |e^{-\delta (t-t^{\prime})}dt^{\prime}.
\end{eqnarray}
After some manipulation of this equation we find:
\begin{eqnarray}
\gamma(t) &=& \Gamma\delta\left |\left[T[t]-T[0]e^{-\delta t}-\delta \int_0^t T(t^{\prime})e^{-\delta (t-t^{\prime})}dt^{\prime}\right]\right |.
\end{eqnarray}
This representation of the rate $\gamma(t)$ has the advantage that it only involves $T(t)$ and does not require the calculation of the derivative of $T$. 
If $T$ exhibits $n$ temperature jumps, one finds:
\begin{eqnarray}
\gamma(t) &=& \Gamma\sum_{i=1}^n \left |\Delta T_i \right |\theta(t-t_i)e^{-\delta(t-t_i)}.
\end{eqnarray}

The ordinary differential equations for the pool model with environmental-shock based back-flow read:
\begin{equation}
    \begin{cases}
    \dot{L} =\gamma(t) G - (\mu + \lambda) L\\
    \dot{G} = \lambda L -\gamma(t) G + \alpha G\left(1-\frac{L+G+D}{N_t}\right)\\
    \dot{D} = \mu  L - \beta D
    \end{cases}
\end{equation}

-  \cmt{Ref to temperature jumps effect. Plots / results.}

>>>>

One of the reasons of the dormant state appearence can be also the temperature shift~\cite{oliver_viable_1995}.
Moreover, the rapid shift in temperature is associated with the appearence of an additional lag phase~\cite{zwietering_modeling_1994}.
To model it we propose a kind of visco-elastic Maxwell type of stress-strain relation and write down the ordinary differential equation for $\gamma$ to be:
\begin{eqnarray}
\dot{\gamma} &=& \Gamma \left |\dot{T}\right |-\delta \gamma.
\end{eqnarray}
We assumed that the direction of the temperature change does not matter, i.e., a change from $T=2$ to $T=12$ has the same effect as a change from $T=14$ to $T=4$. 
$\Gamma$ is a scaling factor and $\delta$ is the relaxation time for the temperature disturbance. 
Integration of this equation yields (with $\gamma(0)=0$):
\begin{eqnarray}
\gamma(t) &=& \Gamma \int_0^t \left |\dot{T}(t^{\prime})\right |e^{-\delta (t-t^{\prime})}dt^{\prime}.
\end{eqnarray}
After some manipulation of this equation we find:
\begin{eqnarray}
\gamma(t) &=& \Gamma\delta\left |\left[T[t]-T[0]e^{-\delta t}-\delta \int_0^t T(t^{\prime})e^{-\delta (t-t^{\prime})}dt^{\prime}\right]\right |.
\end{eqnarray}
This representation of the rate $\gamma(t)$ has the advantage that it only involves $T(t)$ and does not require the calculation of the derivative of $T$. 
If $T$ exhibits $n$ temperature jumps, one finds:
\begin{eqnarray}
\gamma(t) &=& \Gamma\sum_{i=1}^n \left |\Delta T_i \right |\theta(t-t_i)e^{-\delta(t-t_i)}.
\label{eq:gamma_tempshift}
\end{eqnarray}

\begin{figure}[t]
    \begin{center}
    \includegraphics[width=0.67\textwidth]{Figures/pool_model_3pools_resource_tempshift.pdf}
    \caption{{\footnotesize (a) The time dependence of the total bacterial count $N$ for the model including the back-flow to the lag phase due to temperature stres (\ref{eq:ODE_tempshift}).
    (b) The value of the corresponding back-flow rate $\gamma$ calculated according to (\ref{eq:gamma_tempshift}).
    The curves represent no temperature shift (blue), one temperature jump $\Delta T = 10$ at time $t=4$ (red) 
    and the series of different temperature shifts of value 10, 5 and 15 at the tme points 1, 3 asnd 7, respectively (green).
    The paremeters of the back-flow rate are $\Gamma=1$ and $\delta=5$.
    The other model parameters are $\alpha=2.5$, $\mu'=0.5$, $\beta=5$, $\lambda=10^{-2}$, $\mu = 0$ fe each curve.
    }}
    \label{fig:TempJump}
    \end{center}
\end{figure}

The ordinary differential equations for the pool model with environmental-shock based back-flow read:
\begin{equation}
    \begin{cases}
    \dot{L} =\gamma(t) G - (\mu + \lambda) L\\
    \dot{G} = \lambda L -\gamma(t) G + \alpha G\left(1-\frac{L+G+D}{N_t}\right)\\
    \dot{D} = \mu  L - \beta D
    \label{eq:ODE_tempshift}
    \end{cases}
\end{equation}
